---
title: "bio720 assignment 11.18"
author: "Jerry"
date: "2018Äê11ÔÂ18ÈÕ"
output: md_document
---

# Data Visualization with ggplot2: Introduction

## Exploratory vs Explanatory 
- Exploratory visualizations: Graphical data analysis. Easily generated, data-heavy, intended for a small specialist audience.
- Explanatory visualizations: Communication process. Labor-intensive, data-specific, intended for a broader audience.

##  Exploring ggplot2 part1
```{r}
library(ggplot2)
ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point()
```

-Load the ggplot2 package first.

-Deifine x axis and y axis: x = _, y = _)

-geom_point(): Scatter plot

-Notice that ggplot2 treats cyl as a continuous variable, which is wrong.
```{r}
ggplot(mtcars, aes(x = factor(cyl), y = mpg)) +
  geom_point()
```

- Wrapping factor() around cyl: ggplot2 treats cyl as a factor. Problem is fixed.

## Exploring ggplot2 part2
### Other aesthetics
```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()

ggplot(mtcars, aes(x = wt, y = mpg, color = disp)) +
  geom_point()

ggplot(mtcars, aes(x = wt, y = mpg, size = disp)) +
  geom_point()
```

-color =/size = : Legend for the color and size scales was generated(dependent on disp).

### Understanding Variables
ggplot(mtcars, aes(x = wt, y = mpg, shape = disp)) +
  geom_point()

-This(shape = disp) gives you an error: only makes sense with categorical data, and disp is continuous.

## geom_point() and geom_smooth()
```{r}
ggplot(diamonds, aes(x = carat, y = price))  +
  geom_point()

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth()
```

-geom_point(): draw points on the plot.

-geom_smooth(): draw a smoothed line over the points.

## Exploring ggplot2 part4
```{r}
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_smooth()

ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_point(alpha = 0.4)
```

-color = clarity: clarity mapped to color.

-geom_point(alpha = 0.4): This will make the points 60% transparent/40% visible.

## Understanding the grammar part1
```{r}
dia_plot <- ggplot(diamonds, aes(x = carat, y = price))

dia_plot + geom_point()

dia_plot + geom_point(aes(color = clarity))
```

- You can store the plot as a ggplot object that you can use later(add other layers).

- You can also modify the functions.

## Understanding the grammar part2
```{r}
dia_plot + geom_smooth(aes(col = clarity), se = FALSE)
```

-se = FALSE: No error shading.

# Data Visualization with ggplot2: Data

## Base package and ggplot2
-Base package disadvantages: 
 
 1. Plot doesn't get redrawn.
 
 2. Plot is drawn as a image.
 
 3. Need to manually add legend
 
 4. No unified framework for plotting.

-Using ggplot2
 
 The major advantage of ggplot2 is the ggplot object. You can sign the base layers to an object, and then recycle the object with different arguements(such as different plot types).

## Plotting the ggplot2 way
- The ggplot2 way:

Rearrange the data to define variables in your dataset according to what we are really interested in. Should make the function calls as straightforward as possible.

## Tidy Data 
###facet_grid(. ~ ___)
It will generate several plots based on the thing you fill in the blank.

### How to rearrange your data 
-'gather()'
 
 It moves information from the columns to the rows. It takes multiple columns and gathers them into a single column by adding rows. 
 
 Say you have a dataset like this:
 
```{r}
head(iris)
```

Then if you do:

```{r}
library(tidyr)
iris.tidy <- iris %>% 
  gather(key, Value, -Species)
```

And then have a look:

```{r}
 head(iris.tidy)
```

So we can see that all the columns are combined except 'species' column(-species). All the values are now in the 'value' column and the column names in the previous data frame are now combined in the 'key' column.

-'separate()¡®
 
 It splits one column into two or more columns according to a pattern you define.
 So if you add another line to the previous gather command:
```{r}
 iris.tidy <- iris %>%
  gather(key, Value, -Species) %>%
  separate(key, c("Part", "Measure"), "\\.")
```

And then you will see:
 
```{r}
 head(iris.tidy)
```

So the 'key' column in the previous data frame is now separated into "Part" and "Measure". And "\\." lets the 'key' column separate by ".".

-'%>%'
 
 It passes the result of the left-hand side as the first argument of the function on the right-hand side.

-'spread()' 

 It takes two columns (key & value), and spreads into multiple columns: it makes ¡°long¡± data wider.
 
# Data Visualization with ggplot2: Visible Aesthetics

## Typical Aesthetics

X: x axis position

y: Y axis position

colour: Colour of dots, outlines of other shapes

fill: Fill colour

size: Diameter of points, thickness of lines

alpha: Transparency

linetype: Line dash pa!ern

labels: Text on a plot or axes

shape: Shape

### shape

Shapes in R can have a value from 1-25. Shapes 1-20 can only accept a color aesthetic, but shapes 21-25 have both a color and a fill aesthetic.

The default geom_point() uses shape = 19 (a solid circle with an outline the same colour as the inside). 
Good alternatives are shape = 1 (hollow) and shape = 16 (solid, no outline). These all use the col aesthetic.

A really nice alternative is shape = 21 which allows you to use both fill for the inside and col for the outline. 

### hexadecimal colours

Hexadecimal, literally "related to 16", is a base-16 alphanumeric counting system. Individual values come from the ranges 0-9 and A-F. This means there are 256 possible two-digit values (i.e. 00 - FF). Hexadecimal colours use this system to specify a six-digit code for Red, Green and Blue values ("#RRGGBB") of a colour (i.e. Pure blue: "#0000FF", black: "#000000", white: "#FFFFFF"). R can accept hex codes as valid colours.
However, you could use "yellow" as yellow color. "red" as red.

## Modifying Aesthetics
### position = "jitter"

Add some random noise on both x and y to see regions of high density.
It can be used as a argument or a function. 
 
### Scale Function
 
 scale_?_!("_", limits = _, breaks = _, expand = _, labels = _)

 ?: Defines which scale we want to modify(x,y,color...).
 
 !: Match the type of data we are using(continuous or discrette).

 Limits: Scale's limits.

 Breaks: controls the breaks in the guide.

 Expand: is a numeric vector of two, giving a multiplicative and additive constant.
 
 labels: label the categories.

### labs(x =_, y =_)

Quick change the axis labels.
 
# Data Visualization with ggplot2: Geometries
## Scatter Plot

- Each geom has specific aesthetic mappings

- geom_point()

- Essential: x, y

- Optional: alpha, colour, fill, shape, size

-example:
```{r}
 ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, col = Species)) +
 geom_point()
```

-Add layers by '+geom_point()'. Can modify aes inside the brackets.

-crosshair:
 geom_vline(data = _, linetype = _,aes(_intercept = _, col = _))


## Bar Plots
### Histogram
-plot of statistical function

-No space between

-x axis labels between the bars

-example:
```{r}
ggplot(iris, aes(x = Sepal.Width)) +
geom_histogram()
```

-stat_bin: binwidth defaulted to range/30. Use 'binwidth = x' to adjust this:

There will always be a message like this. We can change it anytime.

Different types of 'position':

-stack
```{r}
ggplot(iris, aes(x = Sepal.Width )) +
geom_histogram(binwidth = 0.1,position = "stack")
```

-dodge
```{r}
ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
geom_histogram(binwidth = 0.1, position = "dodge" )
```

-fill
```{r}
ggplot(iris, aes(x = Sepal.Width, fill = Species)) +
geom_histogram(binwidth = 0.1, position = "fill" )
```

### Bar
-geom_bar()

-All positions from before available

-Two types:

Absolute counts

Distributions

-Absolute counts
```{r}
ggplot(sleep, aes(extra)) +
  geom_bar()
```

-Distributions
```{r}
ggplot(mtcars, aes(mpg, fill = cyl)) +
  geom_histogram(binwidth = 1, position = "identity")
```

-geom_errorbar
geom_errorbar(aes(ymin = _, ymax = _),
width = _)

## Line plots
### geom_line
-example:
```{r}
ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_line()
```

For geom_line():

There are several categorical variables:linetype, color, and size are the common choices.

### fill(geom_area)

Instead of overlapping time series, they are added together at each point.

-position = "fill"

For each time point, we have a proportion of the total y value by each category.


### geom_ribbon
ggplot(fish, aes(x = _, y = _, fill = _)) +
geom_ribbon(aes(ymax = _, ymin = 0), alpha = _)